Nondeterministic Bigraphical Reactive Systems for Markov Decision Processes\thanks{Supported by organization x.}

### 2018-08-11 paper.tex
Section 3 is (kind of) done

% TODO: introduce BigraphER \cite{DBLP:conf/cav/SevegnaniC16} and its syntax

\begin{example} \label{example}
  Consider an MDP $(S, \overline{s}, A, P, L)$, where $S = \{ s_0, s_1, s_2, s_3
  \}$, $\overline{s} = s_0$, $A = \{ a, b, c \}$, and $P, L$ defined as follows:
  \begin{equation*}
    \begin{split}
      P(s_0, a) &= [s_1 \mapsto 1], \\
      P(s_1, b) &= [s_0 \mapsto 0.7, s_1 \mapsto 0.3], \\
      P(s_1, c) &= [s_2 \mapsto 0.5, s_3 \mapsto 0.5], \\
      P(s_2, a) &= [s_2 \mapsto 1], \\
      P(s_3, a) &= [s_3 \mapsto 1],
    \end{split}
    \qquad
    \begin{split}
      L(s_0) &= \{ initial \}, \\
      L(s_1) &= \emptyset, \\
      L(s_2) &= \{ heads \}, \\
      L(s_3) &= \{ tails \}.
    \end{split}
  \end{equation*}
  Furthermore, equip it with a reward structure $(\rho, \iota)$, where
  $\rho(s_2) = 3$, $\iota(s_1, b) = 1$, and both functions are zero everywhere
  else.
\end{example}

@@

The MDP can be represented as an NBRS with BigraphER code in
Listing~\ref{example_listing}. BigraphER can construct and visualise the full
transition system, as in Fig.~\ref{example_ts}.

More specifically, reaction rule probabilities are represented as floating-point
numbers inside the arrows (e.g. \texttt{-[0.7]->}), each action encompasses its
reaction rules with \texttt{\textbf{action} actionName} and
\texttt{\textbf{end}}. Rewards can be added to actions simply by inserting an
integer enclosed in squared brackets after the action name (e.g.
\texttt{\textbf{action} b[1]}). Lastly, predicate rewards have the same format,
but are listed on the \texttt{\textbf{preds}} line of the \texttt{\textbf{begin
    nbrs}}---\texttt{\textbf{end}} block (e.g. \texttt{\textbf{preds} =
  \{heads[3]\};}).

Visually, each state is represented by a white ellipse (the starting state is
highlighted with a thicker border), and each action (per state) is a grey
ellipse. The labels inside state ellipses list all predicates that are
satisfied by that state. Edges from actions to states are labelled with both the
probability and the name of the reaction rule. Transitions from either a fully
generated transition system or a simulation, labels from predicates, and
state/transition rewards can be exported to the various plain text formats of a
probabilistic model checker PRISM \cite{DBLP:conf/cav/KwiatkowskaNP11} for
further quantitative analysis.

>>>>

The described extensions have been implemented in an open-source tool for BRSs
BigraphER \cite{DBLP:conf/cav/SevegnaniC16}. We refer the interested reader to
the paper on BigraphER \cite{DBLP:conf/cav/SevegnaniC16}, the related PhD thesis
\cite{DBLP:phd/ethos/Sevegnani12}, and online
resources\footnote{\url{http://www.dcs.gla.ac.uk/~michele/bigrapher.html}} for
further information on the syntax and capabilities of the software. In this
section we will introduce the new features of BigraphER by using a simple MDP as
an example.

\begin{example} \label{example}
  Consider an MDP $(S, \overline{s}, A, P, L)$, where $S = \{ s_0, s_1, s_2, s_3
  \}$, $\overline{s} = s_0$, $A = \{ a, b, c \}$, and $P, L$ are defined as
  follows:
  \begin{equation*}
    \begin{split}
      P(s_0, a) &= [s_1 \mapsto 1], \\
      P(s_1, b) &= [s_0 \mapsto 0.7, s_1 \mapsto 0.3], \\
      P(s_1, c) &= [s_2 \mapsto 0.5, s_3 \mapsto 0.5], \\
      P(s_2, a) &= [s_2 \mapsto 1], \\
      P(s_3, a) &= [s_3 \mapsto 1],
    \end{split}
    \qquad
    \begin{split}
      L(s_0) &= \{ initial \}, \\
      L(s_1) &= \emptyset, \\
      L(s_2) &= \{ heads \}, \\
      L(s_3) &= \{ tails \}.
    \end{split}
  \end{equation*}
  Furthermore, equip it with a reward structure $(\rho, \iota)$, where
  $\rho(s_2) = 3$, $\iota(s_1, b) = 1$, and both functions are zero everywhere
  else.
\end{example}

@@

The MDP can be represented as an NBRS with BigraphER code in
Listing~\ref{example_listing}. More specifically, reaction rule probabilities
are represented as floating-point numbers inside the arrows (e.g.
\texttt{-[0.7]->}), each action encompasses its reaction rules with
\texttt{\textbf{action} actionName} and \texttt{\textbf{end}}. Rewards can be
added to actions simply by inserting an integer enclosed in squared brackets
after the action name (e.g. \texttt{\textbf{action} b[1]}). Lastly, predicate
rewards have the same format, but are listed on the \texttt{\textbf{preds}} line
of the \texttt{\textbf{begin nbrs}}---\texttt{\textbf{end}} block (e.g.
\texttt{\textbf{preds} = \{heads[3]\};}).

BigraphER can construct and visualise the full transition system (see
Fig.~\ref{example_ts}). Each state is represented by a white ellipse (the
starting state is highlighted with a thicker border), and each action (per
state) is a grey ellipse. The labels inside state ellipses list all predicates
that are satisfied by that state. Edges from actions to states are labelled with
both the probability and the name of the reaction rule. Transitions from either
a fully generated transition system or a simulation, labels from predicates, and
state/transition rewards can be exported to the various plain text formats of a
probabilistic model checker PRISM \cite{DBLP:conf/cav/KwiatkowskaNP11} for
further quantitative analysis.
